verisr
========================================================

This package is to support data analysis within the VERIS framework (http://veriscommunity.net).  It is intended to work directly with raw JSON and can be used against the VERIS Community Database (VCDB) found at (http://veriscommunity.net/doku.php?id=public) and (https://github.com/vz-risk/VCDB).

Install it from straight from github:

```{r eval=FALSE}
library("devtools")
install_github("verisr", "vz-risk")
```

To begin, load the package and point it at a directory of JSON files storing VERIS data.

```{r}
library(verisr)
vcdb.dir <- "../VCDB/incidents"
vcdb <- json2veris(vcdb.dir)
```

Now that we have this, we can get a quick view of what's in the data:

```{r}
summary(vcdb)
```

Let's look for a specific variable:

```{r}
ext.variety <- getenum(vcdb, "actor.external.variety")
print(ext.variety)
```

And we could create a barplot with ggplot:

```{r} 
library(ggplot2)
gg <- ggplot(ext.variety, aes(x=enum, y=x))
gg <- gg + geom_bar(stat="identity", fill="steelblue")
gg <- gg + coord_flip() + theme_bw()
print(gg)
```

In progress: searching by two enumerations:

```{r}
hacking.actor <- getenumby(vcdb, "action.hacking.vector", "actor.external.variety")
head(hacking.actor)
```

Now we can create all sorts of views of this data.
For example a faceted bar plot comparing the two:

```{r fig.height=10}
gg <- ggplot(hacking.actor, aes(x=enum, y=x))
gg <- gg + geom_bar(stat="identity", fill="steelblue")
gg <- gg + facet_wrap( ~ primary, ncol=2)
gg <- gg + ylab("Incidents") + xlab("Hacking Vector") 
gg <- gg + ggtitle("External Actors by Hacking Vector")
gg <- gg + coord_flip() + theme_bw()
print(gg)
```

Or perhaps a heat map with the count in the box:
```{r}
gg <- ggplot(hacking.actor, aes(x=enum, y=primary, fill=x, label=x))
gg <- gg + geom_tile() + geom_text()
gg <- gg + scale_fill_gradient(low = "#D8EEFE", high = "#4682B4")
gg <- gg + ylab("External Variety") + xlab("Hacking Vector") 
gg <- gg + ggtitle("External Actors by Hacking Vector")
gg <- gg + theme_bw()
gg <- gg + theme(axis.text.x = element_text(angle = 90, hjust = 1),
                 legend.position="none")
print(gg)